### 2018-10-14 paper/complexity.tex
update of complexity considerations - still incomplete

% -----------------------------------------------------------------------------
\subsubsection*{Process Definition and Basic Properties}
Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
a collection of subsets of $\Sigma$ by
\[
{\cal C} = \{ A\subseteq\Sigma~|~\card{A} = n - \lfloor\frac{n}{2}\rfloor + 1 \}.
\]
%
With this choice of ${\cal C}$, define
\begin{equation}\label{eq:pmax}
\pmax = \Intchoice_{A\in{\cal C}} e:A\then \pmax
\end{equation}
%
The important properties of $\pmax$ are summarised in the following lemma.
%
\begin{lemma}\label{lemma:pmax}
Given alphabet $\Sigma$ with cardinality $\card{\Sigma} = n\ge 2$,
process $\pmax$ fulfils
%
\begin{eqnarray}
{}[\pmax/s]^0 & = & \Sigma \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxa}
\\
\trc(\pmax) & = & \Sigma^*
\label{eq:pmaxb}
\\
\minaccs(P/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxc}
\\
\minhits(P/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxe}
\\
\card{\minhits(P/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor} 
\quad \text{for all $s\in \Sigma^*$}
\label{eq:pmaxd}
\\
\card{H} & = & \lfloor\frac{n}{s}\rfloor \quad \text{for all $H\in {\cal C}$}
\label{eq:pmaxf}
\end{eqnarray}
\end{lemma}
\begin{proof}
 Since  
$
\bigcup_{A\in{\cal C}} A = \Sigma
$
by construction of ${\cal C}$, $[\pmax]^0 = \Sigma$. Since $\pmax/e = \pmax$ for all
$e\in\Sigma$, this proves statements (\ref{eq:pmaxa}) and (\ref{eq:pmaxb}).
The internal choice construct used in the specification of $\pmax$ implies
$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all
$e\in\Sigma$ implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence
of (\ref{eq:pmaxc}).

@@

@@

For the  $U_F(j)$ generated from $\pmax$, the number of test executions
 to be performed is maximal if $\pmax\lessdet_F Q$ and 
$\trc(Q) = \Sigma^*$. In such a situation, no test execution blocks early, because
$Q/s$ can always engage into some $e\in\Sigma$   while $\#s<j$ and will never block
in the last step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is 
offered by the test case. The resulting number of executions is
%
\begin{equation}
\label{eq:maxexec}
\card{\Sigma}^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot \lfloor\frac{n}{2}\rfloor,
\end{equation}
%
because all traces up to length $j$ can be executed with $U_F(j)$ entering branch
(\ref{eq:ufc}), and each of these traces is followed by one event
from each of the hitting sets of ${\cal C}$, and each of these hitting sets 
contains $\lfloor\frac{n}{2}\rfloor$ elements.

@@

>>>>

% -----------------------------------------------------------------------------
\subsubsection*{Process Definition and Basic Properties}
Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
a collection of subsets of $\Sigma$ by
\begin{equation}\label{eq:defC}
{\cal C} = \{ A\subseteq\Sigma~|~\card{A} = n - \lfloor\frac{n}{2}\rfloor + 1 \}.
\end{equation}
%
With this choice of ${\cal C}$, define
\begin{equation}\label{eq:pmax}
\pmax = \Intchoice_{A\in{\cal C}} e:A\then \pmax
\end{equation}
%
The important properties of $\pmax$ are summarised in the following lemma.
%
\begin{lemma}\label{lemma:pmax}
Given alphabet $\Sigma$ with cardinality $\card{\Sigma} = n\ge 2$,
process $\pmax$ fulfils
%
\begin{eqnarray}
{}[\pmax/s]^0 & = & \Sigma \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxa}
\\
\trc(\pmax) & = & \Sigma^*
\label{eq:pmaxb}
\\
\minaccs(P/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxc}
\\
\minhits(P/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxe}
\\
\card{\minhits(P/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor} 
\quad \text{for all $s\in \Sigma^*$}
\label{eq:pmaxd}
\\
\minhits({\cal C})  & = & \{ H\subseteq \Sigma~|~\card{H} = \lfloor\frac{n}{2}\rfloor\}
\label{eq:pmaxf}
\end{eqnarray}
\end{lemma}
\begin{proof}
 Since  
$
\bigcup_{A\in{\cal C}} A = \Sigma
$
by construction of ${\cal C}$, $[\pmax]^0 = \Sigma$. Since $\pmax/e = \pmax$ for all
$e\in\Sigma$, this proves statements (\ref{eq:pmaxa}) and (\ref{eq:pmaxb}).
The internal choice construct used in the specification of $\pmax$ implies
$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all
$e\in\Sigma$ implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence
of (\ref{eq:pmaxc}).

@@

% -----------------------------------------------------------------------------
\subsubsection*{Maximal Number of Test Executions for $\pmax$}

@@

For the  $U_F(j)$ generated from $\pmax$, the number of test executions
 to be performed is maximal if $\pmax\lessdet_F Q$ and 
$\trc(Q) = \Sigma^*$. In such a situation, no test execution blocks early, because
$Q/s$ can always engage into some $e\in\Sigma$   while $\#s<j$ and will never block
in the last step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is 
offered by the test case. The resulting number of executions is
%
\begin{equation}
\label{eq:maxexec}
n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor},
\end{equation}
%
because all traces up to length $j$ can be executed with $U_F(j)$ entering branch
(\ref{eq:ufc}), and each of these traces is followed by one event
from each of the hitting sets of ${\cal C}$.

@@

Summing up  formula 
(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
%
\begin{equation}\label{eq:pmaxcomplexity}
\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  = 
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
\end{equation}
%
as the maximal number of test executions to be performed when testing an error-free
SUT $Q$ against reference process $\pmax$. This maximum is met when $Q$ equals $\Sigma^*$.

% -----------------------------------------------------------------------------
\subsubsection*{Considering Empty Collections of Minimal Hitting Sets}

The argument so far has shown that reference process $\pmax$ requires the most test 
executions among all CSP processes $P$ whose collections of minimal hitting sets  
$\minhits(P/s)$ are never empty for any trace $s$, so that the branch 
(\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.  
Lemma~\ref{lemma:failureshittingsets} implies that an empty collection 
$\minhits(P/s)$
is equivalent to $(s,A)\in\fails(P)$ for all $A\subseteq\Sigma$. This is equivalent to
$\maxrefs(P/s) = \{ \Sigma \}$ or $\minaccs(P/s) = \{ \varnothing\}$.

It remains to 
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider the following process $Z$ whose behaviour is quite close to that of $\pmax$ with the only difference
that after some pre-defined trace $\ol s \in\Sigma^*$ with $\#\ol s < pq$, 
the process has $\Stop$ as
an internal choice, so that $Z/s = \big(\Intchoice_{A\in{\cal C}} e:A\then \pmax\big) \intchoice \Stop$, and, consequently, $\minhits(Z/s) = \varnothing$.

In the following specification of $Z$, given traces $s_1,s_2\in\Sigma^*$, the Boolean 
expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$.
\begin{eqnarray*}
Z & = & Z_1(\varepsilon)
\\
Z_1(s) & = & (s < \ol s)\&\big( \Intchoice_{A\in{\cal C}} e:A\then Z_1(s.e)  \big)
\\ & & \extchoice
\\ & & (s \not< \ol s) \& \pmax
\\ & & \extchoice
\\ & & (s = \ol s) \& \big( \Intchoice_{A\in{\cal C}} e:A\then \max  \big)
\intchoice \Stop
\end{eqnarray*}
%
It is easy to see that $\trc(Z) = \trc(\pmax) =\Sigma^*$ and that for all 
traces $s\neq\ol s$, $\minhits(Z/s) = \minhits(P/s) = \minhits({\cal C})$ as specified
in Lemma~\ref{lemma:pmax}.

When comparing the test cases $U_F(j)$ generated from $\pmax$ and the test cases
$U_F^Z(j)$ created from $Z$, we note that
\begin{enumerate}
\item For $j < \#\ol s$, the test cases  $U_F(j)$ and $U_F^Z(j)$ are identical.
\item For $j = \#\ol s$, branch (\ref{eq:ufb}) can be entered by $U_F^Z(j)/\ol s$, while it is blocked for $U_F(j)/\ol s$, and branch (\ref{eq:ufd}) is blocked
for  $U_F^Z(j)/\ol s$, while it can be entered by $U_F(j)/\ol s$. As a consequence,
the maximal number of executions that can be performed by $U_F^Z(j)$ against the SUT 
is 
\begin{equation}
(n^{j} - 1)\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor} + 1
\end{equation}
according to (\ref{eq:maxexec}): we win one additional execution by entering 
branch (\ref{eq:ufb}) after $\ol s$ and lose $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ executions
by no longer being able to enter branch (\ref{eq:ufd}) after $\ol s$.

\item For $j > \#\ol s$, one execution is won by entering 
branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
\end{enumerate}
%
Summing up over $j = 0,\dots,pq-1$, this results in a difference of 
%
\begin{equation}
(pq  - \#\ol s) - \binom{n}{\lfloor\frac{n}{2}\rfloor}
\end{equation}
executions for the test cases $U_F^Z(j),\ j=0,\dots,pq-1$ in comparison to 
the test cases $U_F(j),\ j=0,\dots,pq-1$. To assess this formula, see the function
table for the first 19  arguments of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ shown
in Table~\ref{table:binom}.
%
\begin{table}[htp]
\caption{Function table of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ for arguments
$n = 2,\dots,20$.}
\begin{center}
\begin{tabular}{c||c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
$n$ & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13  & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\\hline
$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ &
     2 & 3 & 6& 10 & 20 & 35 & 70 & 126 & 252 &
     462&924&1716&3432&6435&12870&24310&48620&92378&184756
\end{tabular}
\end{center}
\label{table:binom}
\end{table}
%
The analysis above leads to the following complexity result.
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then, if $pq < \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
test suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ according to Theorem~\ref{th:failurestest}
requires at most 
%
\begin{equation*}
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
\end{equation*}
%
test executions.
\xbox
\end{theorem}
